---
title: "JSE Project"
output: html_document
---

```{r}
library(dplyr)
SimpsonP <- read.csv(file="Simpson_Stand_Ground_2015.csv",head=TRUE,sep=",")
SimpsonP
```

```{r}
TFull <- xtabs(~Convicted + Accused, data= SimpsonP)
TFull
```

```{r}
SimpsonPVM <- SimpsonP[SimpsonP$MinVictim == 1, ]
head(SimpsonPVM)
TVM <- xtabs(~Convicted + Accused, data= SimpsonPVM)
TVM
VicMinority <- TVM[2, 1] + TVM[2, 2]
Con_Rate_VM <- VicMinority/dim(SimpsonPVM)[1]
Con_Rate_VM

SimpsonPVW <- SimpsonP[SimpsonP$WhiteVictim == 1, ]
head(SimpsonPVW)
TVW <- xtabs(~Convicted + Accused, data= SimpsonPVW)
TVW
VicWhite <- TVW[2, 1] + TVW[2, 2]
Con_Rate_VW <- VicWhite/dim(SimpsonPVW)[1]
Con_Rate_VW
```

The combined conviction rate when the victim was white was  $\frac{`r VicWhite`}{`r dim(SimpsonPVW)[1]`} = `r Con_Rate_VW*100`$ whereas the combined conviction rate when the victim was a minority was  $\frac{`r VicMinority`}{`r dim(SimpsonPVM)[1]`} = `r Con_Rate_VM*100`$. 

```{r}
SimpsonPWD <- SimpsonP[SimpsonP$Accused == "White", ]
#head(SimpsonPWD)
TWD <- xtabs(~WhiteVictim + MinVictim, data= SimpsonPWD)
TWD
WDEF_WV <- TWD[2,1]
WDEF_WV_P <- WDEF_WV/dim(SimpsonPWD)[1]
WDEF_WV_P 

SimpsonPMD <- SimpsonP[SimpsonP$Accused == "Minority", ]
#head(SimpsonPMD)
TMD <- xtabs(~WhiteVictim + MinVictim, data= SimpsonPMD)
TMD
MDEF_MV <- TMD[1, 2]
MDEF_MV_P <- MDEF_MV/dim(SimpsonPMD)[1]
MDEF_MV_P
```

Moreover, most of the white defendants had white victims, 107 of 131, $\frac{`r WDEF_WV`}{`r dim(SimpsonPWD)[1]`}$ = $`r WDEF_WV_P*100`$(81.7%), while most of the minority defendants had minority victims, 64 of 89, $\frac{`r MDEF_MV`}{`r dim(SimpsonPMD)[1]`}$ = $`r MDEF_MV_P*100`$(71.9%). Because
whites defendants mostly had white victims and having a white victim greatly increases the
chance of conviction, at first it appears that the conviction rate is higher among white defendants
because of the defendant’s race.

```{r}
#Convicted <- SimpsonP[SimpsonP$Conviction == "Yes", ]
#WhiteVictim <- SimpsonP[SimpsonP$WhiteVictim == 1, ]
LogRegModel <- glm(Convicted~Accused,family="binomial",data=SimpsonP)
summary(LogRegModel)
LogRegModel$coefficients

MultLogModel <- glm(Convicted~WhiteVictim+ Accused,family="binomial",data=SimpsonP)
summary(MultLogModel)
MultLogModel$coefficients

##1 is row, 2 is column, nothing is the whole thing
A <- prop.table(TFull, 2)
A[1][1]
prop.table(TVM, 2)
prop.table(TVW, 2)


```