---
title: "STAT PR4"
author: 
date: "September 15, 2015"
output: html_document
---

###Investigation 1.6: Improved Baseball Player    

**(b)** State the appropriate null and alternative hypotheses that the manager wants to  to test. [*Hint*: Which is the "by chance" hypothesis and which hypothesis is the manager trying to gather evidence for?]    

*$H_o$: $\pi$ = 0.250*     
*$H_a$: $\pi$ > 0.250*   
    
**(c)** Suppose the manager decides to give the player a sample of 20 at-bats. How likely do you think it is that this 0.333 hitter will be able to convince the manager he has improved from the 0.250 hitter he used to be? Explain.      
```{r}
#E(x) = n*pi
n <-20
pi1 <-0.250
pi2 <-0.333
#probability of showing he is a
Ex1 <-n*pi1 #.250 hitter
Ex2 <-n*pi2 #.333 hitter. 
```
*`r Ex1` for 0.333 hitter and `r Ex2` for 0.250 hitter. By that result, the results' differences are not notable to indicate the improvement.*

So the question is, how likely it is that someone who is now genuinely a 0.333 player will be able to
demonstrate his improvement to the manager (will convince the manager to reject the null hypothesis)?
We will investigate this through a two-step process. First, we need to determine how many hits the
player would need to get to convince the manager that he’s better than a 0.250 hitter. Then we need to
determine how likely a 0.333 hitter is to perform that well.

*There is a notable overlap between the two distributions of 0.25 and 0.333 hitters that it is vague or ambiguous to tell them apart in 20 at-bats. To demonstrate his improvement, the player would have a tough time.* 


**(g)** Based on the simulation results, what is the "empirical level of significance" or the estimated probability that a 0.250 hitter obtains an outcome that is in the rejection region?       
```{r}
l <-1 
while(sum(dbinom(l:20,20,.250))>0.05)
{
  l <-l+1
}

#Find the rejection region.
nullprob <-rbinom(1000, 20, 0.250)
empP <-length(nullprob[nullprob>=l])
```
    
*The empirical probability that the outcome is in the rejection region is `r empP/1000` after the simulation results.*

**(h)** Describe what Type I error means in this context.   
*The manager is convinced with evidences that the player has improved his batting average when the player actually has not improved his batting average.*   

    
**(i)** Step two is to now consider the 0.333 hitter's point of view: What is the empirical probability that he obtains enough hits for the null hypothesis (of him being a 0.250 hitter) to be rejected? In other words, what is the empirical probability that the outcome is in the rejection region? [*Hint*: think about which distribution to look at to answer this question.]   

```{r}

#Finding the rejection region
sample(0:1, 20, replace=TRUE, prob=c(1/3,2/3))  
altprob <-rbinom(1000, 20, 1/3)
empProb <-sum(altprob>=9)
empProb
```  

*The empirical probability that he obtains enough hits for the null hypothesis or the outcome is in the rejection region is $\frac{`r empProb`}{1000}$ $\approx$ `r empProb/1000` after the simulation.* 


**(p)** What might the player ask for in order to have a better chance of showing that his success probability really has improved? Explain. 

*The player should ask for more number of bats or a larger sample size so there would be less overlap in the two sampling distributions to have more power*

**(s)** What is the approximate power of the test for a 0.333 hitter with this cut-off value? How does this approximate power compare to what you found in (i) with a sample size of 20 at-bats?   
```{r}
#Find reject region
n <-1000
i <- 1
while(sum(dbinom(i:100,100,.250))>0.05) #code determines the cut-off value for number of hits
{
  i <-i+1
}
i
```

```{r}
p <- rbinom(n, 100, 0.333)
power <- length(p[p>=i])/n #power or probability of type 2 errors
power
```
*With the increased number of trials, the power is `r power` which definitely improved along with the probability that the hitter would hit the ball.*


**(u)** Now determine the rejection region so that the probability of Type I error (level of significance) is at most 0.01 (still with a sample size of 100 at-bats).   

```{r}
n <- 1000
i <- 1
while (sum(dbinom(i:100,100,.250))>0.01)
  {
  i<-i+1
  }
p <- rbinom(n,100,0.333)
Type2 <- length(p[p>=i])/n
Power <- 1 - Type2
``` 

What is the new cut-off value for the rejection region?   

*The new cut-off value is $\geq$ 36 hits.*

What is the new empirical probability of a Type II error?   

*The new empirical prob of a Type II error is `r Type2`.*

Did this change help the player's likelihood of convincing the manager that he has improved? Explain. 

*As Type I error($\alpha$) decreases, Type II error($\beta$) increases and then power decreases. Therefore, no, this change did not help the player's likelihood of convincing the manager that he has improved*



###RR Chapter 5 Summary   
*This chapter introduces on how to make research's results and processes reproducible through making them available so that anyone can have access to them. There are two available methods for a researcher to store their research in a cloud storage service including Dropbox and GitHub. In order to master using each cloud storage, the researcher needs to master storing, accessing files, collaborating, and versioning control. Cloud storage service provide a sole benefit of allowing the research to be available any where possible in which letting collaborators to  conveniently access, review, comment, and make necessary changes. Moreover, with the aid of version control, it helps researchers to keep track of the work to make easy corrections. I was already aware of GitHub coming from being a Computer Science background, however this chapter did mention the basic of GitHub so that I have a better understanding in using GitHub through R and not Linux environment.*


###Session Info   
```{r}
sessionInfo()
```




